Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM

### 2011-02-21 response.tex

\item \emph{2nd paragraph, The computing power ... Are there references which elucidate this statement?}

A reference was added to a previous study where solving the PNP system in 3D was discussed.

\item \emph{1st sentence, ... principal difficulties ... As mentioned above, there are also other issues
known for the Poisson-Nernst-Planck system in the case of small Debye lenghts. Are the
authors sure that the motivation given in Section is numerical more demanding? If yes, are
there references available?}

In some practical cases such as in case of modeling the Ionic Polymer Metal Composites,
often the values of diffusion constant $D$ and dielectric permittivity $\varepsilon$ are such
that the gradients of $\varphi$ do not become significantly high in a studied time scale, however,
the gradients of $c$ do. This is one reason why it is hard to find an optimal mesh, especially
in case of a 3D domain.

@@

\item \emph{2nd last sentence in 1st paragraph A short explanation, how Hermes further deals with this
difference without going to much into details, might be interesting here ...}

TODO Pavel

\item \emph{1st sentence in Sec. 5.2 Convergence plots of the error, which the authors mention, would be interesting...}

Figure 17 shows an example error convergence in case of HP\_ANISO, HP\_ANISO\_H, and HP\_ISO.

### 2011-02-21 results.tex
Length scale study completed

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.6\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a)
  	and refined mesh (b). The coarse mesh (a)
	and refined mesh (b) were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P).}
  \end{centering}
\end{figure}

@@

\subsection{Length scale study}

The Debye length $\lambda_D$ is the screening length in the electrolyte
solutions. Its numerical value shows the
thickness of the layer in the vicinity of the boundaries $\partial\Omega_1$ 
and $\partial\Omega_2$  where $c \neq 1$.
In all the previous simulations, the Debye screening length was determined by  the constants in
Table~\ref{Table:used-constants} and Eq.~\ref{eq:debye}: $\lambda_D=XXX$. It is known that computation gets increasingly
difficult when reducing the value of $\lambda_D$.
It was our interest to see how small screening lengths can Hermes HP\_ANISO automatic
adaptivity handle. The parameter $\varepsilon$ was varied as follows:
$$\varepsilon_n=\varepsilon\times 0.5^n,$$ where $\varepsilon$ is taken from Table~\ref{Table:used-constants}. The simulations were run for each $\varepsilon_n$ and corresponding $\lambda_D$ value
and maximum number of degrees of freedom and cumulative CPU time were recorded.
The simulation time $t$ for each $\lambda_D$ was chosen to be $\tau$ --- the
characteristic time scale --- and each simulation
was divided equally into fifteen time steps. No PID controller was used.

>>>>

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.9\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	refined mesh (b), and symmetrically refined mesh for length scale study (c). 
	The coarse mesh (a)
	and refined mesh (b) were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P).}
  \end{centering}
\end{figure}

@@

\subsection{Length scale analysis}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{debye_dof}
  \caption{\label{fig:debye_dof} Problem size depending on the
	Debye length. Initially the coarse mesh (shown in Fig.~\ref{fig:mesh}~(a))
	was used. It was necessary to use the fine mesh (see Fig.~\ref{fig:mesh}~(c)) 
	for smaller $\lambda_D$ values.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{debye_cpu}
  \caption{\label{fig:debye_cpu} CPU time depending on the
	Debye length. Initially the coarse mesh (shown in Fig.~\ref{fig:mesh}~(a))
	was used. It was necessary to use the fine mesh (see Fig.~\ref{fig:mesh}~(c)) 
	for smaller $\lambda_D$ values.}
  \end{centering}
\end{figure}
The Debye length $\lambda_D$ is the screening length in the electrolyte
solutions. Its numerical value shows the
thickness of the charged layer in the vicinity of the boundaries $\partial\Omega_1$ 
and $\partial\Omega_2$.
In all the previous simulations, the Debye screening length was determined by  the constants in
Table~\ref{Table:used-constants} and Eq.~\ref{eq:debye}: $\lambda_D=1.7\ \mu m$. It is known that computation gets increasingly
difficult when reducing the value of $\lambda_D$.
It was our interest to see how small screening lengths can Hermes HP\_ANISO automatic
adaptivity handle. The parameter $\varepsilon$ was varied as follows:
$$\varepsilon_n=\varepsilon\times 0.5^n,$$ where $\varepsilon$ is taken from Table~\ref{Table:used-constants}. The simulations were run for each $\varepsilon_n$ and corresponding $\lambda_D$ value
and maximum number of degrees of freedom and cumulative CPU time were recorded.
The simulation time $t$ for each $\lambda_D$ was chosen to be $\tau$ --- the
characteristic time scale --- and each simulation
was divided equally into fifteen time steps. The PID controller was not used.
Fig.~\ref{fig:debye_dof} shows the maximum and average number of degrees of 
freedom during calculation
as a function of the Debye length and Fig.~\ref{fig:debye_cpu} shows cumulative CPU time as a function
of the Debye length. The simulations up to $0.52\ nm$
screening length were carried out on
the initial coarse mesh. However, from $\lambda_D > 0.52\ nm$, the finer initial mesh had to
be used so the existence of the large gradients of the
physical fields $c$ and $\varphi$ near the boundaries could be captured in the first place.
\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{cphitau}
  \caption{\label{fig:cphitau} Calculated fields $c$ and $\varphi$ at $t=\tau=0.81\ ms$
	for $\lambda_D=0.4\ nm$.}
  \end{centering}
\end{figure}
\noindent
The fine mesh allowed simulations with the Debye length down to $0.40\ nm$~m. The calculated
$c$ and $\varphi$ at $t=\tau$ for $\lambda_D=0.40\ nm$ are shown in Fig.~\ref{fig:cphitau}.
It appears that when using even finer initial mesh and higher initial polynomial degrees, even
smaller Debye lengths could be used when necessary. The polynomial space of $c$ had
consistently higher maximum polynomial degree than that of $\varphi$, however, the difference
was less noticeable for smaller Debye lengths.
